Cross-lingual Document Similarity and Event Detection \todo{Better title...}

### 2015-02-23 rupnik2015.tex
- notation changes. small text updates

In short, the article clustering works as follows (details are available in \cite{}). First we create a new document by joining the title and the body of the article. We then tokenize the document, remove the stop words and stem the words. The remaining words are represented as a vector in the vector-space model. Each value in the vector is normalized using the TF-IDF approach, defined as $TFIDF(t,d) = TF(t,d) * IDF(t)$. $TF(t,d)$ represents the \emph{term frequency} (number of times the term $t$ appears in the document $d$), and $IDF(t) = log(N / DF(t))$ is the \emph{inverse document frequency} of term $t$. $IDF(t)$ is computed by comparing the number of documents mentioning the term $t$ ($DF(t)$) with the total number of documents $N$. After normalizing the vector, we use it to find the most similar existing cluster. To be able to compare the vector with a cluster we first need a vector representation for each of the clusters. This is done by computing a centroid vector of the cluster by simply averaging the vectors of the articles in the cluster. By computing the cosine similarity of the article vector with each of the centroid vectors we then find the cluster with the highest similarity. If the highest similarity is above the selected threshold (0.4 in our experiments), the article is assigned to the corresponding cluster, otherwise a new cluster is created, initially containing only the single article.

@@

In order to treat the cluster as describing an event in Event Registry, the cluster has to grow above a certain number of articles (the number depends on the language). This condition ensures that only content that receives a sufficient coverage in the media is treated as an event. Once a new event is identified, a new unique ID is assigned to it and the main information about the event is automatically extracted by analyzing the articles assigned to it. The extracted information includes properties such as the date of the event, the location, who is involved in it, what the event is about, etc.

The clustering of articles is done for each language separately -- English articles are, for example, clustered separately from Spanish articles. When events are identified, they will therefore be identified from clusters in each language separately. This has an undesirable consequence for Event Registry where the same actual event will be represented with two or more event IDs, and each event will only contain coverage in a single language. To avoid this problem we want to exploit the available information about the similarity of the articles in a cluster with articles in other languages. Using this information we hope to be able to connect clusters in different languages and represent them using a single event ID.

@@

Formally, the problem of identifying the clusters describing the same event can be formalized as follows. Each article $a_i$ is written in a language $l$, where $l \in \{A,B,...,H\}$. For each language $l$, a set of clusters $C_l$ is generated and maintained that only contains articles in language $l$. For any cluster $c_i \in C_{l1}$, the task is then to identify other clusters $c_j \in C_{l2}$ (where $l1 \neq l2$) that are describing the same event as $c_i$. We will refer to cluster pairs $c_i$ and $c_j$ that are describing the same event as equivalent clusters. An additional requirement of the task is that identification of the equivalent clusters should not be done exhaustively -- for a cluster $c_i$ we wish to avoid testing all other clusters $c_j$. Performing exhaustive comparison for each cluster $c_i$ would result in $O(N^2)$ tests, where $N$ is the number of all clusters, which is not feasible when the number of clusters is in tens of thousands.

>>>>

In short, the article clustering works as follows (details are available in \cite{}). First we create a new document by joining the title and the body of the article. We then tokenize the document, remove the stop words and stem the words. The remaining words are represented as a vector in the vector-space model. Each value in the vector is additionally normalized using TF-IDF. After normalization, we use the vector to find the most similar existing cluster. To be able to compare the vector with a cluster we first need a vector representation for each of the clusters. This is done by computing a centroid vector of the cluster by simply averaging the vectors of the articles in the cluster. By computing the cosine similarity of the article vector with each of the centroid vectors we then find the cluster with the highest similarity. If the highest similarity is above the selected threshold (we used 0.4 in our experiments), the article is assigned to the corresponding cluster, otherwise a new cluster is created, initially containing only the single article.

@@

In order to treat the cluster as describing an event in Event Registry, the cluster has to grow above a certain number of articles (the number depends on the language). This condition ensures that only content that receives a sufficient coverage in the media is treated as an event. Once a new event is identified, a new unique ID is assigned to it and the main information about the event is then automatically extracted by analyzing the articles assigned to it. The extracted information includes properties such as the date of the event, the location, who is involved in it, what the event is about, etc.

The clustering of articles is done for each language separately -- English articles are, for example, clustered separately from Spanish articles. When events are identified, they will therefore be identified from clusters in each language separately. As a result, the same event will be represented in Event Registry with two or more event IDs, and each event will only contain coverage in a single language. To avoid this problem we want to exploit the available information about the similarity of the articles in a cluster with articles in other languages. Using this information we hope to be able to connect clusters in different languages and represent them using a single event ID.

@@

Formally, the problem of identifying the clusters describing the same event can be formalized as follows. Each article $a_i$ is written in a language $l$, where $l \in L = \{l_1,l_2,...,l_k\}$. For each language $l$, a set of clusters $C_l$ is generated and maintained that only contains articles in language $l$. For any cluster $c_i \in C_{l_a}$, the task is then to identify other clusters $c_j \in C_{l_b}$ (where $l_a \neq l_b$) that are describing the same event as $c_i$. We will refer to cluster pairs $c_i$ and $c_j$ that are describing the same event as equivalent clusters. An additional requirement of the task is that identification of the equivalent clusters should not be done exhaustively -- for a cluster $c_i$ we wish to avoid testing all other clusters $c_j$. Performing exhaustive comparison for each cluster $c_i$ would result in $O(N^2)$ tests, where $N$ is the number of all clusters, which is not feasible when the number of clusters is in tens of thousands.
